---
title: "Frequency of exposure mediates the facilitation effect of cognateness in bilingual lexical acquisition"

pagetitle: Cross-linguistic similarity and  word acquisition

abstract: |
    A prominent feature of the bilingual mental lexicon is that the activation of representations spreads in a language non-selective fashion. An instance of such non-selectivity is embodied by cognateness (i.e., form-similarity between translation equivalents, such as *flower*-*flor* in English and Spanish), which impacts lexical access and processing in bilinguals. Recent studies have suggested that cognateness facilitates vocabulary acquisition in bilingual toddlers---who show larger vocabulary sizes when their languages share many cognates---and an earlier age-of-acquisition for cognates than for non-cognates. The specific mechanisms underpinning such facilitation are unclear. We present an account of bilingual early lexical acquisition in which cognateness interacts with lexical frequency and language exposure to facilitate the acquisition of low-frequency words. We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals aged 12 to 34 months. We used a Bayesian Exploratory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler. We found converging evidence for an earlier age-of-acquisition for cognate words, and for such effect being mediated by lexical frequency and language exposure. Low-frequency words, and words from the language of least exposure were more strongly benefitted by their cognate status than high-frequency words. Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.

thanks: |
    This research was supported by grants from the Spanish Ministerio de Ciencia, Innovación y Universidades (PGC2018-101831-B-I00 and PRE2019-088165), and the Catalan Government [ICREA (Catalan Institution for Research and Advanced Studies) Academia 2019 award]. Gonzalo Garcia-Castro was supported by a fellowship of the Spanish Ministerio de Ciencia, Innovación y Universidades (FPI 2019). The authors declare no conflicts of interest with regard to the funding source of this study. We are grateful to Chiara Santolin, Alicia Franco-Martínez, Cristina Rodríguez-Prada, and Ege E. Özer for their help-ful feedback. We thank Xavier Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical support. We also thank Cristina Dominguez and Katia Pistrin for their efforts in recruiting infants. We would like to thank the clinics Quirón and Sagrada Familia that allowed us to recruit participants in their premises. We also thank all families and infants who participated in the experiments. Data collection was half-way when the COVID-19 pandemic started. We would like to pay special tribute to the families that collaborated with us under these difficult circumstances.

keywords: lexical acquisition, vocabulary, bilingualism, item response theory, bayesian
---


```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load packages
library(ggplot2)
library(patchwork)
library(bayesplot)
library(ggtext)
library(lubridate)
library(brms)
library(glue)
library(dplyr)
library(tidyr)
library(purrr)
library(gt)
library(stringr)
library(tibble)
library(tidybayes)
library(janitor)
library(scales)
library(bvqdev)

# load targets as R objects
tar_load(participants)
tar_load(responses)
tar_load(items)
tar_load(childes)
tar_load(bvq_data)

tar_load(model_fit_4)
tar_load(posterior_draws)

tar_load(model_fit_4_prior)
tar_load(marginal_effects_epreds)
tar_load(posterior_draws_doe)

tar_load(model_log_liks)
tar_load(model_loos)

tar_load(model_rhats)
tar_load(model_neffs)
tar_load(model_mcmc_pairs)
tar_load(model_ppcs)

items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme
bayesplot_theme_set(
    theme_custom()  +
        theme(panel.grid = element_blank(),
              axis.text = element_text(size = 5),
              strip.text = element_text(size = 5),
              plot.title = element_text(size = 8),
              plot.subtitle = element_text(size = 8))
)
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")
color_scheme_set(rep(clrs[4], 6))
options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = scale_color_gradient,
        ggplot2.continuous.colour = scale_color_gradient)



# resolve namespace conflicts
# conflict_prefer("filter", "dplyr")
# conflict_prefer("select", "dplyr")
```


# Introduction {#sec-introduction}

The foundations of word learning are in place early in age. From six months of age, infants start directing their gaze to some objects when hearing their labels, according to both experimental data [@jusczyk1995infants; @tincoff1999beginnings; @bergelson2012months; @bergelson2015early] and parental vocabulary reports [@fenson2007macarthurbates, @samuelson2021precision]. During the last half of their second year, new word forms are acquired at an increasingly fast rate [@goldfield1990early; @fenson1994variability; @mayor2011statistical; @bloom2002how; @bergelson2020comprehension]. Such early stages of lexical acquisition are characterised by substantial variation across children reflected, for instance, on the variability of the number of words they know [i.e. vocabulary size, @fenson1994variability; @frank2021variability]. Despite this variability, trajectories of early vocabulary growth seem quite stable across languages. @tardif2008baby collected data about the first ten words acquired by 10-- to 16--month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese, and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (*daddy*, *mommy*), social routines (*bye*, *uh-oh*) or animals (*woof-woof*). These results were later extended by @frank2021variability to a more diverse set of languages, and found that such cross-language commonalities are stronger at earlier stages of lexical acquisition.

Most literature on early word acquisition still relies extensively on data from monolingual children, and is oblivious to substantial proportion of the world population acquiring more than one language since early ages [@grosjean2021extent]. Previous research about bilingual vocabulary acquisition pointed to bilingual toddlers knowing, on average, less words in each of their languages than their monolinguals peers, but also to both groups knowing a similar amount of words when the bilinguals' two languages are taken into account. @hoff2012dual found that English-Spanish bilingual toddlers in South Florida knew less words in English than monolinguals did. Both groups knew a similar total amount of words when both English and Spanish vocabularies were counted together, which substantiates the importance of collecting data on both languages when assessing bilinguals' communicative development. Other studies have provided converging evidence that bilinguals know a similar or even larger number of words than monolinguals when the two languages are aggregated [@oller2002language; @pearson1994patterns; @pearson1993lexical; @patterson2004comparing; @patterson2004bilingual; @smithson2014bilingualism; @petitto2001bilingual; @gonzalez-barrero2020bilingual]. These studies rely on bilingual samples whose languages are relatively distant in typology, as it is the case of English and Spanish. It is unclear whether children learning typologically more similar languages also know less words in each of their languages than monolinguals. What role could linguistic distance play during early vocabulary growth?

For a given set of concepts, bilingual children may be exposed to two distinct sets of word-forms: one in each language. Depending on the similarity between both languages, the two sets of words may overlap in form in varying degrees. When both languages are typologically close, like Spanish and Catalan (both Romance languages), they are more likely to share a large amount of cognates (i.e., form-similar translation equivalents) than two typologically more distant languages, like Spanish and English (Romance and Germanic languages, respectively). For instance, in the presence of a door, a Spanish-Italian (or a Spanish-Catalan) bilingual might hear *puerta* and *porta* (cognates), whereas a Spanish-English bilingual might hear *puerta* and *door* (non-cognates). It could be the case that mapping two phonologically similar labels (cognates like *puerta*-*porta*) onto the same referent is easier than doing the same with two phonologically dissimilar labels (non-cognates, like *puerta* and *door*). If cognates are easier to acquire than non-cognates, bilinguals learning two languages that share a high proportion of cognates should benefit more often from this facilitation effect than bilinguals learning two languages with a lower proportion of cognates, and should therefore show larger vocabulary sizes.

@floccia2018introduction provided evidence in line with this claim. The authors collected vocabulary data on word comprehension and production from 372 24-month-old bilingual toddlers living in the United Kingdom who were learning English and an additional language. The additional language was one of 13 typologically diverse languages: Bengali, Cantonese Chinese, Dutch, French, German, Greek, Hindi/Urdu, Italian, Mandarin Chinese, Polish, Portuguese, Spanish and Welsh. The authors calculated the average phonological similarity between the words in each of these additional languages and their translation equivalents in English, which was taken as a proxy of the degree of *cognateness* between each pair of languages. Phonological similarity was measured by computing the Levenshtein distance between each cross-language pair of phonological transcriptions. The Levenshtein distance is a metric that computes the distance between two strings of characters by counting the smallest number of insertions, deletions and substitutions one of the strings has to undergo to become identical to the other [@levenshtein1966binary]. For instance, the Levenshtein distance for the translation equivalent *cat*-*kat* (/kæt/-/\textipa{kAt}/) in English and Dutch would be one: a single edition would be needed to make both phonological translation identical (replacing /æ/ by /\textipa{A}/). The same translation equivalent in English and Greek (*cat*-*$\gamma\acute{\alpha}\tau\alpha$*, /\textipa{kæt}/-/\textipa{Gata}/) would score three, as it requires at least three editions: substituting /k/ for /\textipa{G}/, substituting /æ/ for /a/ and adding another /a/. Levenshtein distance scores are then divided by the length of the longest string to normalise the similarity scores between 0 and 1. Following the previous examples: 1/3 (0.33) for /kæt/-/\textipa{kAt}/, and 3/4 (0.75) for /kæt/-/\textipa{Gata}/. Finally, this variable was included as a predictor of participants' vocabulary sizes. Among other findings, Floccia and co-workers reported an increase in productive vocabulary size in the additional language associated with an increase in the average phonological similarity between the translation equivalents of each language pair. For example, English-Dutch bilinguals (22.14% mean cognateness), were able to produce more Dutch words than English-Mandarin bilinguals (1.97% mean cognateness) were able to produce in Mandarin.


Floccia et al. pointed to *parallel activation* as the underpinning mechanism behind their results.  The parallel activation hypothesis stems from the language non-selective account of lexical access, and suggests that bilinguals activate both languages simultaneously during speech production [@costa2000cognate; @hoshino2008cognate] or comprehension [@spivey1999cross; @thierry2007brain], even in fully monolingual environments. One of the clearest pieces of evidence of parallel activation was provided by @costa2000cognate. In this study, Spanish monolinguals and Catalan-Spanish bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (*árbol*-*arbre*, translations of *tree*), whereas in the other half of the trial labels were non-cognates (*mesa*-*taula*, translations of *table*), importantly such distinction was only relevant for bilinguals. Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. In contrast, Spanish monolinguals, who were unfamiliar with the Catalan translations of the Spanish words they uttered, showed equivalent naming times for the two types of stimuli. These results showed that bilinguals' Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures.

Regarding word acquisition, the parallel activation account is a plausible account for Floccia et al.'s results in the light of (1) the evidence pointing to bilingual toddlers' lexicon consisting of a large proportion of translation equivalents: the presence of a word in their lexicon is a good predictor of its translation equivalent also being present [@bilson2015semantic; @tsui2022are; @pearson1995crosslanguage], and that (2) there is experimental evidence of parallel activation during word comprehension in children and toddlers [e.g., @vonholzen2012language; @poulin-dubois2013lexical]. It is possible that cognateness increases the amount of cross-language activation, facilitating word acquisition, and ultimately leading to children learning languages with a larger proportion of cognates (such as children learning English and Dutch) showing larger vocabulary sizes. This account in which cognateness underlies the facilitation effect of typological language distance is also in line with previous studies suggesting that the acquisition of new words is facilitated by their phonological or semantic similarity with other words [@hills2009longitudinal; @fourtassi2020growth; @laing2022phonological; @jones2019children]. If phonological similarity plays a cross-language facilitation role during word acquisition, cognate translation equivalents---which share both semantic and phonological similarity---should be acquired, on average, earlier that non-cognate translation equivalents---which share semantic similarity but not phonological similarity. 

Evidence supporting an earlier age of acquisition for cognates is scarce. @bosch2014first collected parental reports of expressive vocabulary from 48 Catalan-Spanish bilinguals aged 18 months, and found that cognates represented a larger proportion of participant's vocabulary than non-cognates. @schelletter2002effect provided converging evidence from a longitudinal single-case study, in which an English-German bilingual produced cognates earlier than non-cognates, on average. The low sample size in these two studies makes it challenging to draw strong conclusions about the effect of cognateness on vocabulary growth. On the other hand, Floccia et al.'s estimates are statistically more reliable given their remarkably larger sample size (372 participants). Although their results suggest a cognateness effect on vocabulary size, their study was not aimed at testing the effect of cognateness on age of acquisition at the *word-level.* The response variable in this study was the proportion of words each participant understood and/or produced (i.e., vocabulary size), from the list of lexical items in the vocabulary checklists. By aggregating the responses from all words into a single datum per child, information about the acquisition status of cognates vs. non-cognates was no longer available. Also, all participants were aged 24 months, meaning that even if the unaggregated responses to individual items were included as response variable, the possible effect of cognateness could only be interpreted as an increase or decrease in the likelihood of participants at such age to have acquired each item, and not as an increase or decrease in the age of acquisition of such item.

More recently, @mitchell2022cognates addressed this issue providing a more fine grained measure of a  cognateness facilitation effect in vocabulary acquisition. Using a longitudinal sample of 47 16-to-30 month-old French-English bilinguals living in Canada, the authors collected data on expressive vocabulary data in both languages. They created two lists of translation equivalents: one made of 131 cognates, and one made of 406 non-cognates. The proportion of translation equivalents that children were reportedly able to produce was higher in the cognate lists than in the non-cognate list across ages, even when both lists were matched by semantic category (furniture, animals, food were similarity represented in both lists) and age of acquisition norms (an index of word difficulty). These findings shed some light on the ongoing exploration of why (if at all) bilinguals' vocabulary size grows faster when both languages are phonologically more similar. Word-forms sharing more phonemes with their translation equivalents (i.e., cognates) seem to be acquired faster. Parallel activation remains a plausible explanation for this effect. When a child is exposed to a word form, its (phonologically similar) translation equivalent is activated, and this cross-language activation increases the chances of acquisition. The precise underpinnings of this effect are still unclear.

In the present investigation we approach the cognate facilitation on word learning using the conceptual framework of accumulator models of language acquisition [@hidaka2013computational; @mollica2017how]. We build on the assumption that word acquisition is a function of mainly---but not exclusively---the amount of experience a child has accrued with the corresponding word-referent pairing [see @kachergis2022standard for review]. As infants accumulate linguistic input, they also accumulate opportunities to associate familiar word-forms to their corresponding referents. Words with high lexical frequency like 'flower' [`r format(floor(childes[childes$token=="flower",]$freq_million), big.mark = ",")` times per million words in the CHILDES corpus, @macwhinney2000childes] are more likely to be acquired earlier than words with low lexical frequency such as 'headphones' (`r format(floor(childes[childes$token=="taxi",]$freq_million), big.mark = ",")` times per million words) [e.g., @goodman2008does]. Infants might accumulate learning experiences with 'flower' faster than with 'taxi'. In the case of bilinguals, the amount of exposures to a word-form an infants receives is also mediated by their degree of exposure to their languages. Assuming that bilinguals and monolinguals are exposed to an equivalent amount of linguistic input, a Spanish-English bilingual who is exposed to Spanish 80% of the time and to English 20% of the time would need more time to accumulate the same experience with 'flower' and 'taxi' than an English monolingual would (who listens 100% of the time to English).

We suggest that the cognateness effect proposed by @floccia2018results and @mitchell2022cognates facilitates word acquisition by reducing the number of learning instances needed by a word to be acquired, and that this effect takes place *only* once the (cognate) translation of such word has been acquired. We hypothesise that both forms in the translation equivalent (e.g., *flower* and *flor*, for an English-Spanish bilingual) follow their acquisition trajectory in a similar way they would for a monolingual child: accumulating learning instances, which occur a a rate proportional to the lexical frequency of the word-form and the amount of exposure to the language they belong to. When one of the word-forms is sufficiently acquired, the acquisition trajectory of the other is impacted, so that the number of needed learning instances left for acquisition are reduced. This facilitation would occur as the result of the semantic and phonological overlap between both words. When the child has sufficient lexical knowledge of *flower*, and therefore there is some for-referent association in place, future presentations of the similar sounding translation *flor* in the presence of the same reference should lead to an increased likelihood of acquisition. In the case of a non-cognate translation equivalent, such facilitation would be unlikely. For instance, even if the child has already acquired the word *dog*, hearing its dissimilar Spanish translation *perro* would hardly lead to an increased probability of acquisition. In line with this account, only the later-acquired form of the translation equivalent would benefit from its cognate status. Given that children are more likely to acquire words from languages to which they are more often exposed to [@david2008individual; @cattani2014how; @thordardottir2006bilingual], the acquisition of words in the language of lower exposure should, on average, be more susceptible to the effect of cognateness. For instance, if the previous child is exposed 80% of the time to English, and 20% of the time to Spanish, words in Spanish (especially low-frequency ones) would show a cognateness facilitation effect, with an earlier age of acquisition for cognates than for non-cognates.


![Schematic representation of the hypothesised cognate facilitation effect on word acquisition. To the left, the acquisition of a cognate translation equivalent (*flower*-*flor*, which share four phonemes) is affected by the phonological similarity between both word forms. When the child hears *flower*, activation spreads across languages to activate *flor* too. As a result, after exposure to *flower* not only the association between *flower* and its referent is strengthened, but also that of *flower* and its common referent. To the right, the acquisition of a non-cognate translation equivalent (*cow*-*vaca*, which do not share any phoneme) does not undergo the such facilitation. Both word-forms are fairly dissimilar, so that when the child hears *cow*, the activation of *vaca* is unlikely, rendering the effect of phonological similarity negligible.](img/hypotheses.jpg){#fig-hypotheses}

In line with this hypothesis, the size of the effect of linguistic similarity on vocabulary size that Floccia et al. reported was larger in the additional language (language of lower exposure) vocabulary than in English vocabulary (language of higher exposure). Most participants in their sample were English-dominant, meaning that their relative amount of exposure to English was larger than in the additional language. Therefore, participants may have on average learned the English word-form of translation equivalents earlier than the word-form in the additional language. Under this hypothesis, because of children's imbalance in exposure to words in English and in the additional language, the acquisition of English words by English-dominant participants would then benefit less frequently from their cognate status (the other word-form is unlikely to be available yet), while the acquisition of words in the additional language would benefit more frequently from their phonological similarity with the (more likely available) English translation.

We tested the hypothesis that the cognate effect on word acquisition is conditional to a child's exposure rate to the word, which would support the hypothesis that cognateness plays a role in the acquisition of a words *only* once its translation has been already acquired. We collected vocabulary data on production and comprehension from a large sample of bilingual Catalan-Spanish children using an online vocabulary checklist designed for the present study. Participants were aged `r min(floor(participants$age))` to `r max(floor(participants$age))` months, and were exposed to Catalan and/or Spanish to varying degrees. We adopted a Bayesian explanatory item response theory  [*IRT*, see @kachergis2022standard for a similar approach] approach to model the probability of a given participant being reported to *understand* or *understand and say* each word-form in the checklist, conditional to participants age and rate of exposure to the word-form, and the cognate status of the word form, while adjusting for the the number of phonemes in the word form (an indicator of word acquisition difficulty).

We operationalised participants' exposure rate to a word-form using a composite measure that captures the lexical frequency of the word-form (an approximation of how often a participant is exposed to it), and child's amount of exposure to the language such word belongs to. This measure was calculated by multiplying the word-form's lexical frequency with participant's degree of exposure to the language the word belongs to. Lexical frequencies were extracted from the CHILDES corpora [@macwhinney2000childes; @sanchez2019childes] (see Appendix 1 for more details). Participants' degree of exposure were reported by caregivers in a language exposure questionnaire they filled before the vocabulary checklist. Caregivers provided a proportion of the time the child was estimated to listen to any language they were spoken to regularly (described in @sec-methods). The resulting measure is a exposure-weighted lexical frequency measure, which was included in analyses as a proxy of how often bilingual participants were exposed to each word-form.

We defined *cognateness* as a continuous measure of phonological similarity between translation equivalents. We quantified the phonological similarity between each pair of word-forms by computing the inverse of the normalised Levenshtein distance between their X-SAMPA phonological transcriptions [@levenshtein1966binary; @wells1995computer]. This score is computed by counting the number of insertions, deletions and replacements needed by both transcriptions to become identical, dividing the resulting value by the length of the longest transcription, and finally subtracting this value from one. This results in a proportion that indicates how much the two phonological transcriptions of the translation equivalent are similar to each other, ranging from 0% (no similarity at all) to 100% (both transcriptions are identical) [see @floccia2018introduction; @fourtassi2020growth; @laing2022phonological for similar approaches]. For instance, this measure of phonological similarity returns 0% for *ocell* `[uceL]`) and *pájaro* (`[paxa4o]`), Catalan and Spanish for 'bird', and returns 50% for *lluna* (`[Lun5]`) and *luna* (`[luna]`), Catalan and Spanish for 'moon' (see Appendix 1 for more details).

# Methods {#sec-methods}

All materials, data, and reproducible code can be found at the OSF ([https://osf.io/hy984/](https://osf.io/hy984/)) and GitHub ([https://github.com/gongcastro/trajectories](https://github.com/gongcastro/trajectories)) repositories. This study was conducted according to guidelines laid down in the Declaration of Helsinki, and was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I.


## Questionnaire {#sec-questionnaire}

```{r}
#| label: items
#| echo: false
#| message: false
#| warning: false
#| include: false
n_categories <- nrow(distinct(bvq_data$pool, semantic_category))

n_te <- length(unique(bvq_data$pool$te))

n_items <- length(unique(bvq_data$pool$item))

n_item_language <- bvq_data$pool |> 
    count(language) |>
    pull(n) |>
    set_names(c("catalan", "spanish"))

n_items_version <- bvq_data$pool |>
    unnest(version) |> 
    count(version, language) |> 
    pivot_wider(names_from = language,
                values_from = n) |> 
    clean_names() |> 
    mutate(version = tolower(version))

```

To collect vocabulary data from participants, we created an *ad hoc* questionnaire: the Barcelona Vocabulary Questionnaire. This questionnaire was inspired by the MacArthur-Bates Communicative Development Inventory [@fenson2007macarthurbates] and its adaptations to other languages, and was implemented on-line using the formR platform [@arslan2020formr]. This questionnaire is structured in three blocks: (1) a language exposure questionnaire, (2) a demographic survey, and (3) two vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000infant] and consisted in two lists of words: one in Catalan and one in Spanish. both lists included items from a diverse sample of `r n_categories` semantic/functional categories. The Catalan checklist contained `r n_item_language["catalan"]` items and the Spanish checklist contained `r n_item_language["spanish"]`. Items in one language were translation equivalents of the items in the other language, roughly following a one-to-one mapping (e.g., the same participant responded to both *gos* and *perro*, Catalan and Spanish for 'dog').

```{r}
#| label: tbl-items
#| tbl-cap: "Summary of the items included in the final analyses."
#| eval: true
#| echo: false
#| warning: false
items |> 
    arrange(language, list) |> 
    distinct(te, .keep_all = TRUE) |> 
    mutate(a = grepl("A", list),
           b = grepl("B", list),
           c = grepl("C", list),
           d = grepl("D", list)) |> 
    group_by(semantic_category) |> 
    summarise(across(a:d, sum), 
              items = list(meaning),
              .groups = "drop") |> 
    mutate(total = rowSums(across(a:d)),
           items = map(items, function(x) head(x, 2))) |>  
    arrange(desc(total)) |> 
    relocate(a:d, total, items) |> 
    gt(rowname_col = "semantic_category") |> 
    fmt_percent(matches("_prop"), decimals = 1) |> 
    cols_label(a = "List A",
               b = "List B",
               c = "List C",
               d = "List D",
               total = "Total",
               items = "Examples") |> 
    tab_style(cell_text(size = "small"),
              list(cells_body(),
                   cells_stub(),
                   cells_column_labels(),
                   cells_column_spanners(),
                   cells_row_groups())) |> 
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners()) |> 
    tab_style(cell_text(style = "italic"),
              cells_column_labels()) |> 
    tab_options(data_row.padding = px(2),
                row_group.padding = px(1.2)) |> 
    tab_style(cell_text(size = "x-small", align = "left"),
              cells_body(items)) 

```

For each word included in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand *and* say it, or did not understand or say it (checked out by default). Given the large number of words in the vocabulary checklists, we created four different subsets of the complete list of items. Each subset contained a random but representative sub-sample of the items from the complete list. Semantic/functional categories with less than 16 items---thus resulting in less than four items after dividing it in four lists---were not divided in the short version of the questionnaire: all of their items were included in the four lists. Items that were part of the trial lists of some ongoing experiments in the lab were also included in all versions. The resulting reduced vocabulary checklists contained between `r min(n_items_version$catalan)` and  `r max(n_items_version$catalan)` Catalan words, and between `r min(n_items_version$spanish)` and  `r max(n_items_version$spanish)` Spanish words. Participants were randomly allocated into one of four subset. Each response to the questionnaire provided data for a minimum of `r min(n_items_version$catalan, n_items_version$spanish)` and a maximum of `r max(n_items_version$catalan, n_items_version$spanish)` translation equivalents. 

## Participants {#sec-participants}

```{r}
#| label: participants
#| echo: false
#| message: false
#| warning: false
#| include: false
# load packages
n_participants <- responses |>
    distinct(id, age) |>
    count(id) |>
    pull(n) |>
    table()

diff_responses <- responses |>
    distinct(id, age) |>
    left_join(distinct(participants, id, time_stamp, age),
              by = c("id", "age")) |> 
    group_by(id) |> 
    arrange(id, time_stamp) |> 
    mutate(diff_lag = time_stamp-lag(time_stamp, default = NA)) |> 
    summarise(diff_min = min(diff_lag, na.rm = TRUE),
              diff_max = max(diff_lag, na.rm = TRUE),
              .groups = "drop") |> 
    filter(!(diff_min %in% c(-Inf, Inf))) |>
    summarise(diff_min = min(diff_min, na.rm = TRUE),
              diff_max = max(diff_max, na.rm = TRUE))

n_responses <- nrow(distinct(responses, id, age))

dates <- participants$time_stamp |> 
    range(na.rm = TRUE) |> 
    as_date()

participants_edu <- table(participants$edu_parent)
```

We collected `r n_responses` responses to the questionnaire from `r sum(n_participants)` distinct children from the Metropolitan Area of Barcelona between the `r format(dates[1], "%dth of %B, %Y")` and the `r format(dates[2], "%dth of %B, %Y")`: `r n_participants[1]` of those participants participated once, `r n_participants[2]` twice,  `r n_participants[3]` three times, and  `r n_participants[4]` four times. Recurrent participants provided responses with a minimum of `r as.numeric(diff_responses$diff_min)` days between responses, and a maximum of `r as.numeric(diff_responses$diff_max)`, and were always allocated to the same questionnaire list (A, B, C, or D). Participants were part of the database of the Laboratori de Recerca en Infància (Universitat Pompeu Fabra), and were contacted by e-mail or phone if their child were aged between `r floor(min(responses$age))` and `r ceiling(max(responses$age))` months, and had not been reported to be exposed more than 10% of the time to a language other than Spanish or Catalan (see @tbl-participants for a more detailed description of the sample). All families provided informed consent before participating. Upon consent, families were sent a link to the questionnaire via e-mail, which they filled from a computer, laptop, or mobile device. Filling the questionnaire took 30 minutes approximately. After completion, families were rewarded with a token of appreciation.

```{r}
#| label: tbl-participants
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100%, 0%, and 50% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    summary_rows(
        columns = 2:7,
        fns = list("Total" = ~ sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "DoE Catalan") |>
    tab_spanner("Age (months)", 2:8) |>
    tab_footnote(
        "This proportion is complementary to the degree of exposure to Spanish, with the exception of those participants who were also exposed to a third language up to 10% of the time",
        cells_column_labels(columns = doe_catalan)
    ) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

We used the highest self-reported educational attainment of parents or caretakers as a proxy of participants' socio-economic status (SES)This information was provided by each parent or caregiver by selecting one of six possible alternatives in line with the current educational system in Spain: sense escolaritzar/sin escolarizar [no education], educació primària/educación primaria [primary school], educació secundària/educación secundaria [secondary school], batxillerat/bachillerato [complementary studies/high school], cicles formatius/ciclos formativos [vocational training], and educació universitària/educación universitaria [university degree]. Most families reported university studies (`r participants_edu["University"]`, `r percent(participants_edu["University"]/sum(participants_edu))`), followed by families were the highest educational attainment were vocational studies (`r table(participants$edu_parent)["Vocational"]`, `r percent(participants_edu["Vocational"]/sum(participants_edu))`), complementary studies (`r participants_edu["Complementary"]`,  `r percent(participants_edu["Complementary"]/sum(participants_edu))`), secondary education (`r participants_edu["Secondary"]`, `r percent(participants_edu["Secondary"]/sum(participants_edu))`), no formal education (`r participants_edu["No education"]`, <1%) and primary education (`r participants_edu["Primary"]`, <1%).


## Data analysis {#sec-analysis}

```{r}
#| label: sample-sizes
#| echo: false
#| message: false
#| warning: false
#| include: false
n_obs <- nrow(responses)

sum_obs <- responses |>
    mutate(te = factor(te)) |>
    count(te) |>
    summarise(across(n, lst(median, min, max), 
                     .names = "{.fn}"))


# attrition
n_items <- bvq_data$pool

n_items_nouns <- n_items |> 
    filter(class=="Noun") 

n_items_freq <- n_items_nouns |> 
    left_join(childes,
              by = c("childes_lemma" = "token")) |> 
    drop_na(freq_zipf, wordbank_lemma)



n_items_lemma <- n_items_freq |> 
    filter(n_lemmas == 1)

n_items_multi <- n_items_lemma |> 
    filter(!is_multiword) |> 
    filter(include, !is.na(wordbank_lemma)) |> 
    drop_na(sampa) |> 
    distinct(te, language)

n_items_duplicated <- n_items_multi |> 
    add_count(te, name = "n_te") |>
    filter(n_te == 2)

n_items_final <- n_items_duplicated |> 
    count(language) |> 
    group_split(language) |> 
    set_names(c("Catalan", "Spanish")) |>
    map_int("n")

n_te <- length(unique(responses$te))

rope_interval <- percent(c(lower = -0.1/4, upper = 0.1/4), accuracy = 0.1)

r_info <- sessionInfo()

r_version <- glue(r_info$R.version$major, 
                  r_info$R.version$minor, 
                  .sep = ".")

```

From the responses to the `r format(nrow(n_items), big.mark = ",")` items included in the questionnaire, we restricted the analyses to responses to nouns (`r nrow(n_items)-nrow(n_items_nouns)` items corresponding to other grammatical classes were excluded) [@fourtassi2020growth]. We then excluded items with missing lexical frequency scores (*n* = `r nrow(n_items_nouns)-nrow(n_items_freq)`), items that included more than one lemma (e.g., *mono/mico* [monkey], *n* = `r nrow(n_items_freq)-nrow(n_items_lemma)`), multi-word items or phrases (e.g., *barrita de cereales* [cereal bar], *n* = `r nrow(n_items_lemma)-nrow(n_items_multi)`). Finally, we removed items without a translation in the other language (*n* = `r nrow(n_items_multi)-nrow(n_items_duplicated)`). This resulted in a final list of `r sum(n_items_final)` items, corresponding to words `r n_items_final["Catalan"]` Catalan words and their `r n_items_final["Spanish"]` Spanish translations (`r n_te` translation equivalents). After collecting participants responses, the final dataset consisted of `r format(n_obs, big.mark = ",")` observations, each corresponding to a single response of one participant to one item. Each translation equivalent received a median of `r round(sum_obs$median)` responses (*Min* = `r round(sum_obs$min)`, *Max* = `r round(sum_obs$max)`) from participants, both languages pooled together. 

We modelled the probability of participants answering each response category (*No* < *Understands* < *Understands and Says*) using a Bayesian, mixed-effects ordinal regression model (`r dim(model_fit_4$fit)[2]` sampling chains with `r format(dim(model_fit_4$fit)[1], big.mark = ",")` iterations each). This model allowed us to estimate both item and participant word-acquisition trajectories, as a function of different fixed effects from $Age$ (number of months elapsed between participants' birth date and questionnaire completion), $Length$ (number of phonemes in the X-SAMPA
phonological transcription of the word-form), $Exposure$ (composite measure of lexical frequency and language degree of exposure), and $Levenshtein$ (phonological similarity between translation equivalents). For an easier interpretation of the regression coefficients, the values of all predictors were standardised before entering the model by subtracting the mean of the predictor from each value and dividing the result by the standard deviation of the predictor. We also included participant and item random intercepts and slopes when appropriate [@barr2013random]

We compared a baseline model $\mathcal{M}_{0}$, which only included the intercept, $Age$, $Length$, and $Exposure$ as predictors, with increasingly more complex models, introducing predictors in the following order: $Age \times Exposure$ ($\mathcal{M}_{1}$), $Levenshtein$ ($\mathcal{M}_{2}$), $Exposure \times Levenshtein$ ($\mathcal{M}_{3}$), and finally $Age \times Levenshtein$ and $Age \times Exposure \times Levenshtein$ ($\mathcal{M}_{4}$). We compared the predictive performance of the models using Bayesian leave-one-out cross-validation [LOO, @vehtari2017practical]^[Due to computational constraints, we performed LOO based on a random sub-sample of `r format(dim(model_log_liks$model_fit_4)[1], big.mark = ",")` samples of the posterior distribution of the fixed effects [@magnusson2020leaveoneout]], and selected for interpretation the one showing the best performance. More details about the models can be found in Appendix 3.

We assessed the practical relevance of the estimated regression coefficients of the selected model following @kruschke2018bayesian. First, we specified a region of practical equivalence (*ROPE*) from `r rope_interval["lower"]` to +`r rope_interval["upper"]`. This region indicates the range of values that we considered equivalent to zero. The degree of overlap between the posterior distribution of a regression coefficient with the *ROPE* can be interpreted as support for the true parameter of the coefficient being zero or equivalent. Then we computed the 95% highest density interval (*HDI*) of the posterior distribution of each regression coefficient. This is interval contains the true value of this coefficient with 95% probability, given the data. Finally, we calculated the proportion of posterior samples in the 95% *HDI* that fell into the *ROPE*, which indicates the probability that the true value of the regression coefficient falls into the *ROPE*, and should therefore be considered equivalent to zero. For example, a 90% overlap between the *HDI* of a coefficient's posterior distribution and the *ROPE* indicates that, given our data, there is a 90% probability that the true value of the coefficient falls within the *ROPE*, and is therefore equivalent to zero.

Data processing and visualisation was done in R [@rcoreteam2013language, `r r_version`] using the Tidyverse family of packages [@wickham2019welcome]. Bayesian modelling was done using the `brms` [@burkner2017brms; @carpenter2017stan], `loo` [@vehtari2022loo], `tidybayes` [@kay2021tidybayes] and `marginaleffects` [@arel-bundock2022marginaleffects] R packages.

# Results {#sec-results}


```{r}
#| label: results-values
#| echo: false
rhats_df <- map_df(model_rhats, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() 

loo_diff <- loo_compare(model_loos)[2, 1]
loo_diff_se <- loo_compare(model_loos)[2, 2]

post_draws_list <- split(posterior_draws$summary, posterior_draws$summary$.variable)
```


All models showed adequate convergence diagnostics and little evidence of multicollinearity between predictors (see Appendix 4 for details). The extended model ($\mathcal{M}_{4}$), which included the three-way interaction between $Age$, $Exposure$, and $Levenshtein$ showed the best predictive performance, with a difference in expected log-predicted density (*ELPD*) compared to the second best-performing model ($\mathcal{M}_{3}$) of `r round(loo_diff, 2)` $\pm$ `r round(loo_diff_se, 2)`. Given our data, the predictions of the $\mathcal{M}_{4}$ are confidently more accurate than those of the other models (see @tbl-results-loos for all model comparisons). 

```{r}
#| label: tbl-results-loos
#| echo: false
#| message: false
#| warning: false
#| eval: true
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV). The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (model 3)."
model_loos |>
    loo_compare() |>
    as.data.frame() |>
    rownames_to_column("model") |>
    mutate(model = str_remove(model, "model_fit_")) |>
    relocate(model,
             matches("elpd_loo"),
             matches("diff"),
             matches("p_loo"),
             matches("looic")) |>
    gt() |>
    fmt_number(2:9, decimals = 1) |>
    cols_merge_uncert(elpd_loo, se_elpd_loo) |> 
    cols_merge_uncert(p_loo, se_p_loo) |> 
    cols_merge_uncert(looic, se_looic) |> 
    cols_merge_uncert(elpd_diff, se_diff) |> 
    cols_label(
        model = "Model",
        elpd_loo = "ELPD ± SE",
        p_loo = "p ± SE",
        looic = "IC ± SE",
        elpd_diff = "ELPD diff. ± SE"
    ) |>
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) |>
    tab_style(cell_text(weight = "bold"),
              cells_column_labels()) |>
    tab_style(cell_text(align = "left"),
              cells_title(groups = "subtitle"))
```

@tbl-coefs shows the summary of the posterior distribution of the fixed regression coefficients of $\mathcal{M}_{4}$, and their degree of overlap with the ROPE. For interpretability, we report the estimated regression coefficients transformed to the probability scale. The resulting values correspond to the maximum difference in probability of acquisition (*Comprehension* or *Comprehension and Production*) that corresponds to a one standard deviation change in each predictor^[The logit and probability scales relate non-linearly. This means that one logit difference is not necessarily translated to a unique value in the probability scale. For example, the probability of acquisition of a given word might increase in 5% when age increases from 22 to 23 months, the probability of acquisition of the same word might only increase in 0.2% when age increases from 30 to 32 months. The linear growth of the probability of acquisition differs along the logistic curve, and therefore deciding the age point at which to report the estimates of the regression coefficients in the probability scale is not trivial. Following @gelman2020regression, we report the maximum value of such coefficient, which corresponds to the linear growth (i.e. derivative) of the logistic curve at the age at which most participants were acquiring a given word. This value can be approximated by dividing the coefficient in the logit scale by four: $\hat{\beta_j}/4$, where $\hat{\beta_j}$ is the estimated mean of the posterior distribution of coefficient $j$.].

```{r}
#| label: tbl-coefs
#| tbl-cap: "Posterior distribution of regression coefficients in the probability scale. *Median*: median of the posterior distribution in the logit scale. *95% HDI*: 95% highest density interval (*HDI*) of the distribution; narrowest range of values that contains 95% of the distribution, and therefore is 95% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
posterior_draws$summary |>
    select(.variable_name, .median, .lower, .upper, .rope_overlap, .type) |>
    mutate(
        across(.median:.upper,
               function(x) {
                   ifelse(
                       .type=="Intercepts (at 22 months)",
                       plogis(x),
                       x/4
                   )
               }
        )) |> 
    gt(groupname_col = ".type", 
       rowname_col = ".variable_name") |>
    fmt_percent(c(.median, .rope_overlap)) |>
    fmt_number(.lower:.upper, scale_by = 100) |> 
    cols_merge(c(.lower, .upper),
               pattern = "[{1}, {2}]") |>
    cols_label(
        .variable_name = "Parameter",
        .median = "Median",
        .lower = md("95% *HDI*"),
        .rope_overlap = md("*ROPE* prob.")
    ) |>
    tab_style(cell_text(weight = "bold"),
              cells_column_labels(columns = 1:5)) |>
    tab_style(cell_text(align = "left"),
              cells_title(groups = "subtitle")) |> 
    tab_style(cell_text(size = "small"), 
              list(cells_body(),
                   cells_row_groups(),
                   cells_stub())) |> 
    tab_style(cell_text(style = "italic"),
              cells_row_groups())
```


The word exposure index ($Exposure$) had a strong effect on the probability of acquisition ($\beta$ = `r percent(post_draws_list[["b_exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std"]]$.upper, accuracy = 0.01)`]). All of the posterior samples of this regression coefficient excluded the *ROPE*. The impact of this predictor on the probability of acquisition was positive: for every standard deviation increase in exposure, the participant was `r percent(post_draws_list[["b_exposure_std"]]$.median/sd(responses$exposure), accuracy = 0.01)` more likely to acquire it.

The effect of phonological similarity by itself (as indicated by the regression coefficient of the main effect of $Levenshtein$) was equivalent to zero ($\beta$ = `r percent(post_draws_list[["b_lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_lv_std"]]$.upper, accuracy = 0.01)`]), as `r percent(post_draws_list[["b_lv_std"]]$.rope_overlap, accuracy = 0.01)` of the posterior samples of its regression coefficient fell into the *ROPE*.

However, the 95% *HDI* of the regression coefficient of the $Exposure \times Levenshtein$ interaction excluded the *ROPE* entirely ($\beta$ = `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the effect of phonological similarity on a word's probability of acquisition changed depending on participants' exposure to word. Follow-up analyses on this interaction (see @fig-results-marginal) showed that, when exposure to word was low (e.g., -1 SD), phonological similarity increased the probability of acquisition substantially. This effect was negligible when for words with median or high exposure (+1 SD).

The 95% *HDI* of the regression coefficient of the $Age \times Exposure$ interaction did not exclude the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std"]]$.upper, accuracy = 0.01)`]): `r percent(post_draws_list[["b_age_std:exposure_std"]]$.rope_overlap, accuracy = 0.01)` of its posterior samples fell within the *ROPE*. This shows that the effect of the word exposure index on the speed at which participants acquired words across ages was inconclusive: given our data, it is difficult to tell whether $Exposure$ increased or decreased the slope of the words' acquisition trajectories.

The 95% *HDI* of the regression coefficient of the $Age \times Levenshtein$ interaction overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:lv_std"]]$.upper, accuracy = 0.01)`]), indicating that phonological similarity did not speed-up the acquisition trajectories of the words across ages. 

Finally, the 95% *HDI* of the regression coefficient of the $Age \times DoE \times Levenshtein$ interaction also overlapped completely with the *ROPE* ($\beta$ = `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.median, accuracy = 0.01)`, 95% *HDI* = [`r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.lower, accuracy = 0.01)`, `r percent(post_draws_list[["b_age_std:exposure_std:lv_std"]]$.upper, accuracy = 0.01)`]), suggesting that the impact of the word exposure index on the effect of phonological similarity did not differ substantially across ages. 

```{r}
#| label: fig-results-marginal
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| eval: true
#| fig-cap: "Expected posterior predictions. Posterior predictions for the extended model were generated by computing the probability acquisition that resulted from feeding the equation of the regression model with values of the posterior distribution of each coefficient. These values correspond to samples of the posterior distribution that were drawn during the MCMC estimation of the model. We generated 25 posterior predictions for a series of combination of levels of interest. Each thin line and corresponds to one prediction for one set of predictor values. Thicker lines indicate the median posterior prediction. The X-axis indicates the age (in months) for which the prediction is generated. The Y-axis indicates the predicted probability of acquisition. Different colours indicate different levels of phonological similarity (as indicated by the Levenshtein similarity between pairs of translation equivalents). Predictions are presented separately for different degrees of word exposure index: little exposure to the word, mean exposure, and high exposure. Predictions for *Comprehension* are show on top and predictions for *Comprehension and Production* are shown on the bottom. In-sample predictions lie inside the grey rectangles."

clrs_reorder <- c("#58508d", "#003f5c", "#122A36", "#ffa600", "#ff6361", "#bc5090")

marginal_effects_epreds |> 
    mutate(
        age = rescale_variable(age_std, 
                               mean(responses$age), 
                               sd(responses$age)),
        exposure = factor(
            exposure_std,
            levels = unique(exposure_std),
            labels = c("Low exposure (-1 SD)",
                       "Mean exposure",
                       "High exposure (+1 SD)"),
            ordered = TRUE
        ),
        lv = factor(
            lv_std,
            levels = unique(lv_std),
            labels = paste0(c(0, 50, 100), "% similarity"),
            ordered = TRUE
        ),
        .category = ifelse(.category=="Understands",
                           "Comprehension",
                           "Production")
    ) |> 
    ggplot(aes(age, .value, colour = lv, fill = lv)) +
    facet_grid(.category~exposure) +
    # stat_summary(fun.data = median_hdi, geom = "ribbon",
    #              alpha = 0.5, colour = NA) +
    annotate(geom = "rect",
             xmin = min(responses$age),
             xmax = max(responses$age),
             ymin = -Inf,
             ymax = Inf,
             fill = "grey90",
             colour = NA) +
    annotate(geom = "text",
             x = min(responses$age),
             y = 1,
             hjust = -0.075,
             vjust = 1,
             label = "Observed\nage range",
             colour = "grey40",
             size = 2) +
    geom_vline(xintercept = mean(responses$age), colour = "grey",
               linewidth = 1) +
    geom_hline(yintercept = 0.5,
               colour = "grey",
               linewidth = 1) +
    geom_line(aes(group = interaction(lv, .draw)),
              linewidth = 0.75, alpha = 0.15) +
    stat_summary(fun = median, geom = "line", linewidth = 1) +
    labs(
        x = "Age (months)",
        y = "Probability of acquisition",
        linetype = "Phonological similarity (Levenshtein)",
        colour = "Phonological similarity (Levenshtein)",
        fill = "Phonological similarity (Levenshtein)",
        subtitle = glue("**Phonological similarity**<br> <span style = 'color:{clrs_reorder[5]};'>0% similarity</span>   <span style = 'color:{clrs_reorder[4]};'>50% similarity</span>   <span style = 'color:{clrs_reorder[1]};'>100% similarity</span>")) +
    scale_colour_manual(values = clrs_reorder[c(5, 4, 1)]) +
    scale_fill_manual(values = clrs_reorder[c(5, 4, 1)]) +
    scale_y_continuous(
        breaks = seq(0, 1, 0.25),
        limits = c(0, 1),
        labels = percent
    ) +
    scale_x_continuous(breaks = seq(5, 50, 5),
                       limits = c(10, 50)) +
    theme(
        legend.position = "none",
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, 
                                    colour = "black", 
                                    linewidth = 0.75),
        plot.subtitle = element_textbox_simple(
            fill = "grey90",
            padding = margin(8, 8, 8, 8),
            margin = margin(0, 0, 10, 0)
        ))
```

# Discussion {#sec-discussion}

In this study, we investigated the role of cognateness on word acquisition. We contrasted two hypotheses regarding the word acquisition process at which the possible (facilitative) effect of cognateness might take place. First, we considered the possibility that cognateness boosts acquisition at early stages, before a stable, robust representation of either of the word-forms involved in the translation equivalent has been formed in the child's lexicon. This account predicts that the effect of cognateness would take place early in age, and would be symmetric across languages: one word-form boosts the acquisition of the other via their phonological similarity. If this hypotheses were supported by data, we would expect to observe and early difference in probability of comprehension and/or production between more similar and less similar translation equivalents (i.e., an effect of $Levenshtein$ in interaction with $Age$), with this difference being equivalent for the word-for to which the child is exposed to more frequently, and for the word-form to which the while is exposed less frequently (no interaction effect between $Exposure$ and $Levenshtein$).

We also considered an alternative hypothesis: that the facilitation effect of cognateness takes place at later stages of word acquisition, when at least one of the word-forms of the translation equivalent hold a detailed, robust representation in the child's lexicon. Only then, the phonological similarity between the acquired word-form and its yet-to-be-acquired translation would boost the acquisition of the latter. This account predicts a larger effect of cognateness for word-forms to which the child is exposed to less frequently, and therefore are likely to be acquired later than their translation. This effect of cognateness would also be expected to take place at later ages. Data backing this account would support a model in which the effect of $Levenshtein$ interacts with $Exposure$, and in which possibly this interaction, or at least the effect of $Levenshtein$ interacts with age.

To test these hypotheses, we used an *ad-hoc* vocabulary checklist to collect parental reports of word acquisition status from children learning Catalan and/or Spanish, who were exposed to varying degrees of exposure to either language. We then used item response theory (*IRT*) to estimate the probability of each word-form being acquired by a child---either as comprehension or production---as a function of the child's $Age$, their $Exposure$ to the word-form (operationalised as a composite measure of lexical frequency and language exposure), and the phonological similarity of the word-form with its translation in the other language ($Levenshtein$). We also added the $Length$  of the word-form as a nuance predictor the increase the predictive performance of the model.

Our results are, to a large extent, in line with previous studies. As it could be expected, we found that shorter word-forms were acquired at early ages than longer word-forms. For instance, $sol$ (Catalan for *sun*) was acquired earlier than *cocodril* (Catalan for *crocodile*). Shorter word-forms consist on fewer phonemes and, frequently, fewer syllables. This makes them simpler to encode as phonological representations in the lexicon, and therefore easier to process. Ultimately, these words are easier to produce, and therefore to be reported as acquired in parental vocabulary checklists. Previous studies found word length to be more strongly associated with receptive vocabulary than with expressive vocabulary [e.g., @braginsky2019consistency; @jones2019children]. 

According to previous literature, we also anticipated that words to which children were more frequently exposed would be acquired earlier [e.g., @braginsky2019consistency]. In line with accumulator models of word acquisition, the age of acquisition of a word is a function of the child's rate of exposure to the word, which conspires with child- and word-level characteristics to the formation of a representation of the word in the child's lexicon [@mollica2017how; @hidaka2013computational]. The amount of learning instances that a child is given to learn a word is given, to a large extent, by the lexical frequency of such word in the child's language. Words with higher lexical frequency are more likely to be uttered in the presence of the child, while words with lower lexical frequency appear less often. In the case of bilinguals, the exposure rate to a given word not only depends on lexical frequency, but also on the child's amount of exposure to the language such word belongs to [@cattani2014how; @floccia2018results]. For instance, a Spanish dominant bilingual child who is exposed to Catalan only 10% of the time will rarely hear the Catalan word *casa*, despite its high lexical frequency. We estimated the exposure rate of the words in our list by calculating a language exposure-weighted measure of lexical frequency, which adjusts the lexical frequency of the word by the child's amount of exposure to the language such word belongs to. We found a strong association between this measure of exposure rate and word's age of acquisition. Interestingly, this predictor had a stronger role during earlier ages: younger infants benefited from higher exposure rates at earlier ages, suggesting that early word acquisition is supported by mere exposure to word forms more strongly than later word acquisition.

Finally, we found a facilitation effect of cognateness, providing converging evidence with previous studies [@floccia2018results; @mitchell2022cognates; @bosch2014first; @schelletter2002effect]. Critically, this effect was mediated by exposure: only words to which children were exposed to less frequently benefited from their cognate status. These results favour the account that a word-form benefits from its phonological similarity with its translation only after such translation has been acquired first. On the one hand, our results provide converging evidence for @floccia2018results, and arguably account for the language distance facilitation effect they found in the additional (non-English) language of their bilingual sample. Most children in their sample were English-dominant, and therefore their exposure to English was higher than to the additional language. However, it was in the additional language vocabulary where they found a facilitation effect of cognateness. Bilingual children learning two languages that shared many cognateness (e.g., English-German) had larger vocabularies in their additional language than those bilingual who were learning two languages that shared less cognates (e.g., English-Chinese). Given our results, we argue that it is language exposure that mediates the relationship between cognateness and vocabulary size in @floccia2018results's sample. 


In line with the hypothesis that received more support from our data, we predicted an interaction effect between participants' age, cognateness, and word exposure, since we expected cognateness to play a more important role at later stages of word learning, when at least one of the members of the translation equivalent has been acquired, and therefore its counterpart in the other language can benefit from its cognate status. We found evidence supporting that the true value of this interaction is close to zero. Finding such an effect would provide further evidence that cognateness benefits word acquisition only after the translation of such word is acquired. However, the absence of such interaction does not preclude a more careful endorsement
of this hypothesis. It is possible that, even if the cognateness effect plays a role if the acquisition of the word-form with the least exposure of the translation equivalent, age itself might not modulate the size of this effect: cognateness plays a similar role across ages.

This study has several limitations. One of them is the fact that the languages involved are typologically very close. Catalan and Spanish are both Romance languages, share a considerable amount of cognates, and very often co-exist in the same social spaces (families, schools, working place, etc.). Although there are no populations of bilinguals with identical sociolinguistic characteristics, and therefore it would be pointless to compare the sample in this study to any "standard" population of bilinguals, the generalisability of our results should be taken with a grain of salt. It is possible, for instance, that the large amount of cognates that Catalan-Spanish bilingual children are exposed leads them to learn at early stages of word acquisition that words tend to sound similar in both languages, and therefore to be more keen on assuming that similar sounding words are semantically related. Children exposed to a pair of languages sharing a fewer amount of cognates might not make the same assumption, even when the word-forms they hear in the presence of the same referents are very similar at the phonological level.

Another limitation is the predominantly cross-sectional design of our data set. We collected vocabulary data from children aged 12 to 32 months, a critical period in development in which the foundations of a lexicon get established. We estimated word acquisition curves by modelling the probability of comprehension and production across ages. However, most participants provided data at only one age point. Effectively, we are making longitudinal inferences based on cross-sectional data, and drawing conclusions about the development of a single participant across multiple age points using data from different participants at individual age points. For this reason, and also because of the observational nature of our study, causal inferences derived from our analyses are unwarranted [@kraemer2000how; @kesmodel2018crosssectional]. One counterpoint to this shortcoming is that, while data our data is cross-sectional if participants are considered as the observational unit, data about word-forms/translation equivalents is longitudinal. The same translation equivalent is responded to at multiple ages. Since translation equivalents are our main observational unit of interest over which we primarily conduct statistical inference (e.g., estimating the size of the cognateness-effect), this provides some further justification for our conclusions. 

Future studies addressing this subject may benefit for a more adequate design oriented towards making causal inferences. One such design can be partially satisfied by (1) collecting a more substantial longitudinal dataset, in which the same participant provides data points for each translation equivalent at multiple ages, and (2) testing the role of cognateness and word exposure in an experimental context, in which such variables can be orthogonally manipulated or randomised. This second possibility would also allow to explore an interpretation of the results obtained in the context of *word learning*, which is not possible in the present study. Another consideration for future investigations is the involvement of a more diverse range of language pairs as in @braginsky2019consistency or @floccia2018introduction to improve the generalisability of the resulting inferences across more populations of bilinguals.

## Conclusions

In summary, our study has replicated previous studies and provided novel insights about word acquisition in bilingual contexts, and how the presence of cognates in the children's linguistic input impacts the early formation of the lexicon. We found that during the acquisition of low frequency words, bilingual children seem to benefit more strongly from the word's phonological similarity with its translation in the other language. We suggest that this effect might be explained by the need of a robust lexical representation of one of the translations for cognateness to play a role.


# References {#sec-references}

::: {#refs}

:::

{{< pagebreak >}}

# Appendix 1: model predictors {.appendix}

{{< include _model-predictors.qmd >}}

{{< pagebreak >}}

# Appendix 2: model details {.appendix}

{{< include _model-details.qmd >}}

{{< pagebreak >}}

# Appendix 3: convergence diagnostics {.appendix}

{{< include _model-diagnostics.qmd >}}

{{< pagebreak >}}


# Appendix 4: frequency and language exposure as separate predictors {.appendix}

{{< include _model-doe.qmd >}}

{{< pagebreak >}}

